State-Trait Models

### 2024-05-22 examples/_queue/_11_Examples_MultivariateLatentGrowthCurveModels.qmd
update examples 10-14

---
title: "Autoregressive Crosslagged Panel Models"
author: "Biostatistics Working Group"
---

@@

## Overview
Autoregressive Cross-Lagged Panel Analysis (ARCL) is a statistical technique used to examine the reciprocal relationships between variables over time. This method incorporates both autoregressive paths, which account for the stability of each variable across time points, and cross-lagged paths, which assess the influence of one variable on another over subsequent time points.

This example will examine the height and weight across 3 annual measurement occasions in a sample of youth taking part in the ABCD Study. The primary aim is to explore the reciprocal relationships between height and weight at different time points, analyzing how height at one time point predicts weight at future time points (and vice-versa). ARCL will be used to model both the stability of height and weight over time (autoregressive paths) and the potential cross-lagged influences between consecutive measurements of these two constructs.

## Preliminary Setup
### Load R libraries
```{r}

library(tidyverse) # Collection of R packages for data science
library(gtsummary) # Publication-ready tables
library(rstatix) # Statistical Tests in tidy format
library(lavaan) # Structural Equation Modeling in R
library(semPlot) #

@@

## Results
### Compute Autoregressive Crosslagged Panel Analysis
This code performs an Autoregressive Cross-Lagged Panel Analysis (ARCL) to examine the reciprocal relationships between 'Height' and 'Weight' measurements across 3 annual visits. The model output is provided below, along with a summary of results. This approach models both the stability of height and weight over time (autoregressive paths) and the potential cross-lagged effects between these constructs.

```{r}

# Reshape data from long to wide format
df_wide <- df_long %>%
    pivot_wider(
        id_cols = c(id),
        names_from = event,
        values_from = c(height, weight),
        names_sep = "_"
    )

@@

model <- "
  # Autoregressive paths
  weight_Year_3 ~ a2 * weight_Year_2
  weight_Year_2 ~ a1 * weight_Year_1
  weight_Year_1 ~ a2 * weight_Baseline
  height_Year_3 ~ a2 * height_Year_2
  height_Year_2 ~ a1 * height_Year_1
  height_Year_1 ~ a2 * height_Baseline

# Cross-lagged paths
height_Year_3 ~ b1 * weight_Year_2
height_Year_2 ~ b1 * weight_Year_1
height_Year_1 ~ b2 * weight_Baseline
weight_Year_3 ~ b1 * height_Year_2
weight_Year_2 ~ b1 * height_Year_1
weight_Year_1 ~ b2 * height_Baseline
"

# Fit the model to your data (replace 'your_data' with the name of your data frame)
fit <- sem(model, data = df_wide)

# View the results
summary(fit)

@@

These results show significant autoregressive effects for both height and weight, indicating that measurements at each time point strongly predict subsequent measurements (e.g., height at Year 2 significantly predicts height at Year 3 with an estimate of 1.05, p < .001). The cross-lagged effects indicate a small but significant negative influence of weight on height at later time points (e.g., weight at Year 2 negatively affects height at Year 3 with an estimate of -0.007, p < .001). The covariances between height and weight at Year 3 are also significant, suggesting a relationship between these variables at the final time point.

@@

# Fit and plot the model with adjusted label positions and curved edges
# Assume 'fit' is your fitted SEM model
semPlotObj <- semPaths(
    fit,
    whatLabels = "est", # Show estimated coefficients
    edge.label.cex = 0.6, # Smaller font size for edge labels
    layout = "tree",
    rotation = 2,
    edge.label.position = 0.75, # Moves labels closer to the destination node
    edge.curve = 0.3, # Increase curvature to separate labels from the path
)

# Adding a title using base R graphics functions
title(main = "Autoregressive Cross-Lagged Panel Model for Height and Weight")

# Redraw the plot with potentially adjusted parameters for better visibility
print(semPlotObj)

@@

This plot illustrates the autoregressive cross-lagged panel model for height and weight across 3 annual measurement occasions. The model includes both autoregressive paths (predicting a variable from its previous time point) and cross-lagged paths (predicting a variable from the other variable at a previous time point).

The autoregressive paths show that height and weight at each time point strongly predict their respective values at subsequent time points. For example, height at Year 2 significantly predicts height at Year 3. These paths are indicated by curved lines connecting each time point for height and weight, with coefficients showing the strength of these predictions.

The cross-lagged paths reveal how height and weight influence each other over time. For instance, weight at Year 2 negatively affects height at Year 3, as indicated by the coefficients on the curved paths connecting weight at one time point to height at the next.

The overall model highlights the strong stability of height and weight over time (autoregressive effects) and the modest influence of weight on height and vice versa (cross-lagged effects). This plot helps visualize the dynamic interplay between height and weight, providing insights into how these variables evolve together over time.

## Wrapping Up
The autoregressive cross-lagged panel model analysis revealed significant relationships between children's height and weight measured over 3 annual assessments. The model showed strong autoregressive effects, indicating that height and weight measurements at each time point strongly predict their subsequent measurements. For example, height at Year 2 significantly predicts height at Year 3 with an estimate of 1.05 (p < .001).

Additionally, the cross-lagged effects revealed a small but significant negative influence of weight on height at later time points. Specifically, weight at Year 2 negatively affects height at Year 3 with an estimate of -0.007 (p < .001). The significant covariances between height and weight at Year 3 suggest a relationship between these variables at the final time point.

The autoregressive cross-lagged panel model framework used in this example demonstrated its utility in capturing both the stability of measurements over time and the dynamic interplay between height and weight. This approach is particularly useful in developmental studies, as it allows for the examination of reciprocal influences between variables over time, providing a comprehensive understanding of the growth trajectories and their interdependencies.

More generally, this method is valuable for understanding both individual-specific trajectories and the relationships between multiple developmental indicators in longitudinal data. By accounting for the autoregressive and cross-lagged effects, this model offers a nuanced perspective on how different aspects of growth interact and influence each other over time.

>>>>

---
title: "Multivariate Latent Growth Curve Models"
author: "Biostatistics Working Group"
---

@@

## Overview

In this example, we will utilize xxxxxx models to analyze xxxxxx obtained across multiple measurement occasions for a sample of youth participating in the ABCD Study. Our primary objective is to understand xxxxxx, while factoring in the clustered nature of observations within individuals over time. XXXX models facilitate this by xxxxx.

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables
library(lavaan)       #
```

@@

## Results
### Compute Multivariate Latent Growth Curve Model
This code performs multivariate latent growth curve to examine the reciprocal relationships between 'Height' and 'Weight' measurements across 3 annual visits. The model output is provided below, along with a summary of results. This approach xxxxx.

```{r}
# Reshape data from long to wide format
df_wide <- df_long %>%
    pivot_wider(
        id_cols = c(id),
        names_from = event,
        values_from = c(height, weight),
        names_sep = "_"
    )

@@

# Load necessary library
library(lavaan)

# Specify the multivariate latent growth curve model
model <- '
  # Intercept and slope factors for variable 1 (height)
  i_var1 =~ 1*height_Baseline + 1*height_Year_1 + 1*height_Year_2
  s_var1 =~ 0*height_Baseline + 1*height_Year_1 + 2*height_Year_2

  # Intercept and slope factors for variable 2 (weight)
  i_var2 =~ 1*weight_Baseline + 1*weight_Year_1 + 1*weight_Year_2
  s_var2 =~ 0*weight_Baseline + 1*weight_Year_1 + 2*weight_Year_2
'

# Fit the model using FIML for handling missing data
fit <- sem(model, data = df_wide, missing = "fiml")

# Check the summary to identify potential issues
summary(fit, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)

@@

These results show xxxxx, indicating that xxxxx (e.g., xxxxx). The xxxx effects indicate xxxx.

@@

@@

This plot illustrates xxxx.

## Wrapping Up
The multivariate latent growth curve revealed xxxxxxxx.

### 2024-05-22 examples/_queue/_12_Examples_LatentTransitionAnalysis.qmd
update examples 10-14

---
title: "Multivariate Latent Growth Curve Models"
author: "Biostatistics Working Group"
---

@@

@@

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables
library(lavaan)       #
```

@@

## Results
### Compute Multivariate Latent Growth Curve Model
This code performs multivariate latent growth curve to examine the reciprocal relationships between 'Height' and 'Weight' measurements across 3 annual visits. The model output is provided below, along with a summary of results. This approach xxxxx.

```{r}
# Reshape data from long to wide format
df_wide <- df_long %>%
    pivot_wider(
        id_cols = c(id),
        names_from = event,
        values_from = c(height, weight),
        names_sep = "_"
    )

df_wide <- na.omit(df_wide)

# Load necessary library
library(lavaan)

# Specify the multivariate latent growth curve model
model <- '
  # Intercept and slope factors for variable 1 (height)
  i_var1 =~ 1*height_Baseline + 1*height_Year_1 + 1*height_Year_2
  s_var1 =~ 0*height_Baseline + 1*height_Year_1 + 2*height_Year_2

  # Intercept and slope factors for variable 2 (weight)
  i_var2 =~ 1*weight_Baseline + 1*weight_Year_1 + 1*weight_Year_2
  s_var2 =~ 0*weight_Baseline + 1*weight_Year_1 + 2*weight_Year_2
'

# Fit the model using FIML for handling missing data
fit <- sem(model, data = df_wide, missing = "fiml")

# Check the summary to identify potential issues
summary(fit, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)

>>>>

---
title: "Latent Transition Analysis"
author: "Biostatistics Working Group"
---

@@

XXXXXXXXXXX Models xxxxxxx. This xxxxxxx.

@@

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables
library(tidyLPA)      # XXXXXXXX

@@

## Results 
### Compute Latent Transition Analysis
This code performs latent transition analysis to examine the reciprocal relationships between 'Height' and 'Weight' measurements across 3 annual visits. The model output is provided below, along with a summary of results. This approach xxxxx.

```{r}
# Estimate the LTA model
model <- tidyLPA::estimate_profiles(df_wide, 2, variances = "equal", covariances = "zero")

summary(model, fit.measures=TRUE, standardized=TRUE, rsquare=TRUE)

plot(model)

### 2024-05-22 examples/_queue/_13_Examples_GrowthMixtureModels.qmd
update examples 10-14

# XXXXXXXXXXX {.unnumbered}

@@

<p>

@@

</p>

## Preliminary Setup

::: panel-tabset
### Install Packages {.tabset .tabset-fade .tabset-pills}

::: blue
> **This code installs the r packages necessary for this example, if
> they are not already installed**

```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

# Create a list of required packages
packages_required <- c("tidyverse","rstatix","DT")

# Check which packages are not installed and install them
packages_to_install <- setdiff(packages_required, rownames(installed.packages()))
if (length(packages_to_install) > 0) {
    install.packages(packages_to_install)
}

# Load the required packages
lapply(packages_required, library, character.only = TRUE)

```
:::

### Load Packages

::: blue
**This code loads the r libraries necessary for this example**

```{r lib-load}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables

```
:::

### Config Options

::: blue
**This code configures knitr code chunk options**

```{r config}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::
:::

## Descriptives Overview

::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

::: blue
**This code reads in and shows the data to be used in the current
example**

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname", "nihtbx_totalcomp_fc", "anthroweightcalc", "anthroheightcalc")
subset_data <- data_nonimaging[, selected_vars]

library(dplyr)
# # Merge the datasets on 'src_subject_id' and 'eventname'
merged_data <- data_demographics %>%
  full_join(subset_data, by = c("src_subject_id", "eventname"))

# Inspect the merged data structure
str(merged_data)

# Define event names to be retained in the analysis and convert variables to appropriate data types
eventnames_to_include <- c("baseline_year_1_arm_1",
                           "1_year_follow_up_y_arm_1",
                           "2_year_follow_up_y_arm_1",
                           "3_year_follow_up_y_arm_1",
                           "4_year_follow_up_y_arm_1")

df <- merged_data %>%
  filter(eventname %in% eventnames_to_include) %>%
  mutate(
    src_subject_id = as.factor(src_subject_id),
    eventname = factor(eventname, levels = eventnames_to_include, ordered = TRUE),
    age = as.numeric(age),
    sex = as.factor(sex),
    race.4level = as.factor(race.4level),
    hisp = as.factor(hisp),
    high.educ.bl = as.factor(high.educ.bl),
    household.income.bl = as.factor(household.income.bl),
    acs_raked_propensity_score = as.numeric(acs_raked_propensity_score),
    rel_family_id.bl = as.factor(rel_family_id.bl),
    site_id_l = as.factor(site_id_l),
    nihtbx_totalcomp_fc = as.numeric(nihtbx_totalcomp_fc),
    anthroweightcalc = as.numeric(anthroweightcalc),
    anthroheightcalc = as.numeric(anthroheightcalc)
  ) %>%
  # Exclude cases from unused assessment waves
  filter(!is.na(eventname))

```
:::

### Descriptives

::: blue
**This code creates a descriptives table**

```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

# Define a function to compute descriptives
compute_descriptives <- function(data, event_name) {
  # For factor variables
  sex_desc <- paste0(table(data$sex), " (", round(100 * prop.table(table(data$sex)), 1), "%)")
  race_desc <- paste0(table(data$race.4level), " (", round(100 * prop.table(table(data$race.4level)), 1), "%)")

  # For numeric variables
  age_desc <- paste0(round(mean(data$age, na.rm = TRUE), 2), " (", round(sd(data$age, na.rm = TRUE), 2), ")")
  weight_desc <- paste0(round(mean(data$anthroweightcalc, na.rm = TRUE), 2), " (", round(sd(data$anthroweightcalc, na.rm = TRUE), 2), ")")
  height_desc <- paste0(round(mean(data$anthroheightcalc, na.rm = TRUE), 2), " (", round(sd(data$anthroheightcalc, na.rm = TRUE), 2), ")")

  # Combine into a data frame
  desc_df <- data.frame(
    Variable = c("Sex - Female", "Sex - Male or other", "Race - Asian", "Race - Black", "Race - Other/Mixed", "Race - White", "Age", "Weight", "Height"),
    Value = c(sex_desc, race_desc, age_desc, weight_desc, height_desc)
  )

  # Rename the Value column based on event name
  colnames(desc_df)[2] <- event_name
  return(desc_df)
}

# Compute descriptives for each event
baseline_desc <- compute_descriptives(subset(df, eventname == "baseline_year_1_arm_1"), "baseline_year_1_arm_1")
one_year_desc <- compute_descriptives(subset(df, eventname == "1_year_follow_up_y_arm_1"), "1_year_follow_up_y_arm_1")
two_year_desc <- compute_descriptives(subset(df, eventname == "2_year_follow_up_y_arm_1"), "2_year_follow_up_y_arm_1")
three_year_desc <- compute_descriptives(subset(df, eventname == "3_year_follow_up_y_arm_1"), "3_year_follow_up_y_arm_1")

# Join all data frames side-by-side
final_table <- baseline_desc %>%
  left_join(one_year_desc, by = "Variable") %>%
  left_join(two_year_desc, by = "Variable") %>%
  left_join(three_year_desc, by = "Variable")

# Adjust for the required format
final_table[1:6, 3:5] <- NA

# Round numeric values to two decimal places
numeric_cols <- sapply(final_table, is.numeric)
final_table[numeric_cols] <- lapply(final_table[numeric_cols], round, 2)

# Create heading rows with the same columns as final_table
heading_rows <- data.frame(
  Variable = c("Sex", "Race"),
  baseline_year_1_arm_1 = NA_real_,
  `1_year_follow_up_y_arm_1` = NA_real_,
  `2_year_follow_up_y_arm_1` = NA_real_,
  `3_year_follow_up_y_arm_1` = NA_real_
)

# Set column names of heading_rows to match final_table
colnames(heading_rows) <- colnames(final_table)

# Introduce group labels and adjust the "Variable" column
final_table <- rbind(
                    heading_rows[1,], 
                    final_table[1:2,],
                    heading_rows[2,], 
                    final_table[3:6,], 
                    final_table[7:9,]
                    )

# Update the Variable column to remove redundant factor variable name
final_table$Variable <- gsub("Sex - ", "", final_table$Variable)
final_table$Variable <- gsub("Race - ", "", final_table$Variable)
final_table$Variable[final_table$Variable == "Male or other"] <- "Male"

# Add non-breaking spaces for increased indentation
final_table$Variable[final_table$Variable %in% c("Female", "Male", "Asian", "Black", "Other/Mixed", "White")] <- 
  paste0(rep("&nbsp;", 6), final_table$Variable[final_table$Variable %in% c("Female", "Male", "Asian", "Black", "Other/Mixed", "White")])

# Update column headers
colnames(final_table)[2:5] <- c("Baseline", "Year 1", "Year 2", "Year 3")

# Display the table interactively without row names, with updated column headers, and with HTML entities rendered
datatable(final_table, 
          colnames = c("", "Baseline", "Year 1", "Year 2", "Year 3"),
          options = list(pageLength = nrow(final_table), autoWidth = TRUE),
          rownames = FALSE, escape = FALSE,
          caption = "Descriptives Table")  # Add table title

@@

:::
:::

## Results 
<div class = "blue">

::: panel-tabset
### Build Model {.tabset .tabset-fade .tabset-pills}

::: blue

**STEP 1: XXXXXXXXX **

----

```{r}
# Combine the data from all classes
data <- rbind(class1, class2, class3)
```

### Model Specification and Estimation
Now that we have our simulated data, we can fit a Growth Mixture Model. In this example, we will assume that there are three latent classes. However, in practice, the number of classes is usually unknown and must be determined through model comparisons or other methods.
```{r, eval=FALSE}
# Fit the Growth Mixture Model
gmm <- hlme(value ~ time,
            random = ~ time,
            subject = "id",
            mixture = ~ time,
            data = data,
            ng = 3,
            B = list(value = c(1, 0, 0, 0)
```

```{r}
# Estimate the model
fit <- lavaan::sem(model, data = df_wide, missing = "FIML")

@@

>>>>

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered

# This code configures knitr code chunk options

knitr::opts_chunk$set(
    echo = T, message = F, warning = F, error = F, output = T,
    comment = NA, cache = T, code_folding = T,
    R.options = list(width = 220, digits = 3),
    fig.align = "center",
    out.width = "75%", fig.asp = .75
)
```

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

# This code loads the data to be used for the example

library(tidyverse)

# Set the data path
data_path <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
df <- readRDS(data_path)

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, anthroheightcalc, anthroweightcalc) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(anthroheightcalc, anthroweightcalc) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        anthroheightcalc = round(as.numeric(anthroheightcalc), 2),
        anthroweightcalc = round(as.numeric(anthroweightcalc), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = anthroheightcalc,
        weight = anthroweightcalc
    )

df_long <- df_long %>%
    group_by(id) %>%
    filter(all(c("Baseline", "Year_1", "Year_2", "Year_3") %in% event)) %>%
    ungroup()

@@

@@

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(lavaan)       #
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables

@@

### Descriptives
```{r}
descriptives_table <- df_long %>%
    select(event, height, weight) %>%
    mutate(event = factor(event)) %>%
    tbl_summary(
        by = event,
        missing = "no",
        label = list(height ~ "Height", weight ~ "Weight"),
        statistic = list(all_continuous() ~ "{mean} ({sd}) )", all_categorical() ~ "{p}%"),
    ) %>%
    modify_header(all_stat_cols() ~ "**{level}**<br>N = {n}") %>%
    bold_labels() %>%
    italicize_levels() %>%
    modify_spanning_header(all_stat_cols() ~ "**Assessment Wave**")
theme_gtsummary_compact()

descriptives_table
```

## Results 
### Compute Growth Mixture Model
This code performs growth mixture modeling to examine the reciprocal relationships between 'Height' and 'Weight' measurements across 3 annual visits. The model output is provided below, along with a summary of results. This approach xxxxx.

```{r}
# Combine the data from all classes
#data <- rbind(class1, class2, class3)

# Fit the Growth Mixture Model
gmm <- hlme(value ~ time,
            random = ~ time,
            subject = "id",
            mixture = ~ time,
            data = data,
            ng = 3,
            B = list(value = c(1, 0, 0, 0)

# Estimate the model
fit <- lavaan::sem(model, data = df_wide, missing = "FIML")

@@

These results show xxxxx, indicating that xxxxx (e.g., xxxxx). The xxxx effects indicate xxxx.

### Model Plots
```{r}

This plot illustrates xxxx.

## Wrapping Up
The multivariate latent growth curve revealed xxxxxxxx.

### 2024-05-22 examples/_queue/_14_Examples_StateTraitModels.qmd
update examples 10-14

---
title: "Growth Mixture Models"
author: "Biostatistics Working Group"
---

@@

## Results 
### Compute Growth Mixture Model
This code performs growth mixture modeling to examine the reciprocal relationships between 'Height' and 'Weight' measurements across 3 annual visits. The model output is provided below, along with a summary of results. This approach xxxxx.

```{r}
# Combine the data from all classes
#data <- rbind(class1, class2, class3)

# Fit the Growth Mixture Model
gmm <- hlme(value ~ time,
            random = ~ time,
            subject = "id",
            mixture = ~ time,
            data = data,
            ng = 3,
            B = list(value = c(1, 0, 0, 0)

# Estimate the model
fit <- lavaan::sem(model, data = df_wide, missing = "FIML")

@@

## Wrapping Up
The multivariate latent growth curve revealed xxxxxxxx.

>>>>

---
title: "State-Trait Models"
author: "Biostatistics Working Group"
---

@@

## Results 
### Compute Latent State-Trait Model
This code performs latent state-trait analysis to examine xxxxx. The model output is provided below, along with a summary of results. This approach xxxxx.

```{r}

# Reshape data from long to wide format
df_wide <- df_long %>%
    pivot_wider(
        id_cols = c(id),
        names_from = event,
        values_from = c(height, weight),
        names_sep = "_"
    )

df_wide <- na.omit(df_wide)

# Define the model
model <- '
  # Trait component
  trait =~ 1 * anthroheightcalc_0 + 1 * anthroheightcalc_1 + 1 * anthroheightcalc_2 + 1 * anthroheightcalc_3

  # State component
  state =~ anthroheightcalc_0 + anthroheightcalc_1 + anthroheightcalc_2 + anthroheightcalc_3

  # Residual variances
  anthroheightcalc_0 ~~ anthroheightcalc_0
  anthroheightcalc_1 ~~ anthroheightcalc_1
  anthroheightcalc_2 ~~ anthroheightcalc_2
  anthroheightcalc_3 ~~ anthroheightcalc_3
'

# Estimate the State-Trait Model
fit <- sem(model, data = df_wide, missing = "FIML")

@@

## Wrapping Up
The latent state-trait analysis revealed xxxxxxxx.
